---
title: "Practice Activity Week 2"
format: html
embed-resources: true
editor: source
---

1. Write a function called `summ_stats()` that takes a vector `x` as
input and produces the following output:

- for **numeric variables**: returns the mean, median, standard deviation, and
IQR as a data frame
- for **categorical variables**: returns the number of levels (categories) of
the variable

*Hint:* You can use `tibble()` to create the data frame.
    * Example: `tibble(a = 1:2, b = 2:3)` creates a data frame with variables
    `a` and `b`.

```{r}
#| label: sum-stats-function
library(dplyr)
summ_stats<- function(x){
  if(is.numeric(x)){
    summ<-tibble(mean=mean(x),med=median(x), sd=sd(x),iqr=IQR(x))
    return(summ)
  }
  else if (is.factor(x)||is.character(x)){
    numLevels<-nlevels(x)
    return(numLevels)
  }
}

tib<- tibble(a = 1:2, b = 2:3)
summ_stats(tib$a)

```

2. Confirm that your function works on each type of variable by running the 
code below. 

```{r}
library(ggplot2)
summ_stats(diamonds$carat)

summ_stats(diamonds$color)
```

3. Use `map()` to apply your `summ_stats()` function to every column in the 
`diamonds` dataset. 
*Hint 1: Look up the `bind_rows()` documentation from dplyr to combine summary statistics for all quantitative variables into one data frame.*
*Hint 2: The `.id` argument will be especially helpful in adding the variable names!*


```{r}
#| label: sum-stats-map
library(purrr)
stats<-diamonds %>% select(where(is.numeric)) %>%   
  map(summ_stats) %>%
  bind_rows(.id = "variable")
stats
```

4. Let's make the output more intuitive, with the variables on the columns and
the summary statistics on the rows. 

*Hint:* You will need to do a double pivot (`pivot_longer()` then
`pivot_wider()`) to achieve this result!

```{r}
#| label: double-pivot
library(tidyr)
stats |>
  pivot_longer(c(mean,med,sd,iqr), names_to = "Statistics",values_to="values")|>
  pivot_wider(names_from = variable,values_from = values )

```

## Required for STAT 541, Optional for STAT 431

5.  Now that we have the output we want, let's use our code to write a
`summ_df()` function that takes a data frame as an input and outputs a table of
summary statistics for every variable in the data frame.
*Hint: The body of your function should contain __all__ the code from Question 4.*

```{r}
#| label: df-function

```

6. Demonstrate that your function works using the `mpg` dataset. 

```{r}
#| label: try-df-function


```
